---
title: "Homework 6"
output: 
  html_document: 
    keep_md: yes
---

Welcome to homework 6! For this homework, I've been asked to choose from a list of 6 tasks. I have chosen to do tasks 2 and 6. I'll first be creating my own function and will finish up by exploring nested dataframes.

Prompt:
"Write a couple of experimental functions exploring some functionality that is useful to you in real life and that might form the basis of your personal package."

Because I work in the social sciences, I find myself doing a LOT of copy/pasting ggplot2 code for histograms and barplots. These figures constitute 75% of my "exploratory data analysis," so it would be very useful if I could automate this process. My goal then is to create a function that will produce histograms of every variable in any dataframe that I read into R. Afterwards, I would like to find a way to automate the generation of 2X2 barplots to show interation effects between two or more independent variables and a single dependent variable. Perhaps later this can serve as the basis for an R package of my creation. 

Some people might suggest plotting several histograms at once (in a single panel) using various R packages, but I find this clunky. I like being able to look at each variable in isolation. 

First, I'll read in some data. This is a dataset called EWB light that was collected by personality psychologists.

```{r}
ewb <- read.table("C:/Users/Wade/Dropbox/Psych 546E 2016/Datasets/ewblight.txt", header=TRUE)
head(ewb)
```

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(psych))
# Psych is an R package developed for psychologists. It's mostly used to print tables of 
# summary statistics that are commonly used in psychology.
```

For loops follow the following syntax:

for (var in seq) {
  expr
}

That is to say: For each variable in this sequence (or each element in this set), execute this expression.

You can also insert a break statement if you don't want the for loop to be performed on the entire sequence. Ex:

for (var in seq) {
  if (nchar(var) > 10) {
  break
  }
  expr(var)
}
 
You can use "next" instead of break if you want to skip a variable. 

```{r}
names(ewb)
```


```{r}
# for(name in names) {
#  hist(ewb[,name])
# }
```

This code appears to work, but the titles are all messed up. I have to use a for loop with a different syntax to put the column names on the graphs.

```{r}
for (i in 1:11) {
  hist(ewb[,name],
       main = paste(colnames(ewb)[i]),
       xlab = "") 
}
```

Perfect! Now I have code that will generate histograms for every variable in a dataset. It may not look like it, but finding the right code for this took me a surprisingly long time.

Next, let's try to implement ggplot2 instead of just using base R. We'll explore this in the context of 2X2 interaction bar plots.

```{r}
for (i in 1:11) {
  hist(ewb[,name],
       main = paste(colnames(ewb)[i]),
       xlab = "") 
}
```




```{r}
?paste 
(paste(ewb$name))
paste(colnames(ewb))
```







<h4>Process</h4>
I had to do some online research into how for loops work.

